Web Processing - Standardised GIS Analyses for Cable Route Planning

### 2023-02-05 paper/paper-CableRoute.tex
added missing requirements, typos and types
linear regression of estimation time and total costs
refactor paper-CableRoute.tex to exclude the larger sections results and methods.
These were moved into there own files

\newacronym{wfs}{wfs}{Web Feature Service}
\newacronym{wps}{wps}{Web Processing Service}
\newacronym{res}{res}{resolution}
\newacronym{r}{r}{ratio}
\newacronym{al}{al}{all touched}
\newacronym{t}{t}{True}
\newacronym{f}{f}{False}
\newacronym{mmd}{mmd}{mean minimum distance}
\newacronym{agg}{agg}{aggregated}
\newacronym{rel}{rel}{relative}
\newacronym{corr}{corr}{corrected}

@@

	\section{Methods}
	We retrieve a set of different spacial data-sets from  public sources as a basis to create the cost raster.
	Field of study are the counties of Cuxhaven and Osterholz in the state of Lower Saxony, Germany.
	Areas protected by different European and National conservation laws are provided by the German Environment Agency
	as \acrfull{wfs}~\cite{noauthor_schutzgebiete_2015}.

	The nation wide land coverage (ATKIS) with a scale of 1:250000 are provided by the Federal Agency for Cartography
	and Geodesy\cite{noauthor_digitales_2021}.
	The nation wide power grid (tags: 'power': line) has been retrieved via OpenStreetMap.\cite{boeing_osmnx_2017}
	Local data as houses at Level of Detail 1 are offered by the State Office for Geoinformation and Land Surveying of
	Lower Saxony\cite{noauthor_opengeodatani_2022}.
	In addition local planning geodata for the land usage are taken from
	from 'Metropolplaner' (Planing data Lower Saxony \& Bremen)\cite{noauthor_metropolplaner_2022}

	PyWPS\cite{noauthor_welcome_2016} is used to offer the least cost path algorithm as a \acrfull{wps}.
	The for the initial implementation of the least cost path algorithm the implementation for the QGIS-Plugin
	'Least Cost Path'\cite{noauthor_leastcostpathdijkstra_algorithmpy_2022} has been taken into account.

	The different layers from the different entities are optionally filtered, buffered and than rasterized.
	Filtering the layers of the files for special attributes enables to further differentiate further.
	For examples makes it possible to differentiate between heath and uncultivated land.
	Adding a buffer can be used either used to convert a line objects as power grids and streets into a polygon with the
	correct physical width, or to add minimum distance from an existing of planed area to the new power grid.
 	Each of theses rasterized layers are given a weight, or cost that expresses the cost of using land covered by this layer.
 	The costs of all layers of is aggregated with the maximum function.
	Thus, an area covered by multiple layers is uniformly used with the highest costs.
	For every area in the study area, that is not covered by any layer, is given the default cost.
	The costs has been grouped into different levels~(see table~\ref{tab:1}) starting from preferential areas with
	very low costs, via no restrictions, which is the default, used when no other layers are covering the local area,
	to restricted, strongly restricted and prohibited areas with high costs.
	These higher costs resemble the degree how much a local area should be avoided, while routing the path.
	The ratio of the higher costs to the lower costs directly translates into the additional diversion in pixels
	the algorithm is willing to go, for avoiding an area of high costs.
	Thus, as prohibited areas describe a legal obligation, not to use these areas or only to the utmost minimum,
	the weight that resembles the costs for these types of areas, has to be especially high. \\
	The rasterization transforms a vector into a raster.
	The rasterization can be executed in to different ways.
	In both ways, the rasterization can be imaged, as the old vector is layered over the new raster grid with the new
	given resolution and the new affine transformation and the coordinate reference system of the vector.
	Both rasterization techniques differ  in who the pixels are selected, that describe, that a geo-object overlays the pixel.
	The pixels can either be described as overlaid with a geo-object, when either the center of that pixel is overlaid
	by the geo-object, or any part of the pixel is overlaid.
	Hence, the version with any part overlaid is called all touched True.
	The version where a overlaid pixel center is require is thus called all touched False.
	All touched False is considered the default (see figure~\ref{fig:alltouched}).
	\begin{figure}[!ht]
		\centering

		\resizebox{8cm}{!}{
			\begin{tikzpicture}
				\node[draw] at (0,6.5) {\huge all touched false};
				\draw[step=1cm,gray,very thin] (-1.9,-1.9) grid (5.9,5.9);
				\draw[fill=blue, thick] (-0.2, 0.2)--(-0.6, 3.1)--(-0.2,4.1)--(1.2,3.4)--(2.8,3.6)--(3.2, 2.6)--(3.4, 2.1)--(1.8,1.3)--(1.2, 0.1)--cycle;

				\draw[ultra thick,->, red] (6.5, 2) --node[above=1mm] {\huge rasterization} (8.5, 2);

				\draw[step=1cm,gray,very thin] (9.1,-1.9) grid (16.9,5.9);
				\draw[fill=blue] (11,0)--(11,2)--(10,2)--(10,3)--(11,3)--(11,4)--(12,4)--(12,3)--(13,3)--(13,4)--(14,4)--(14,2)--(13,2)--(13,1)--(12,1)--(12,0)--cycle;
			\end{tikzpicture}%
		}%
		\vspace*{5mm}
		\resizebox{8cm}{!}{
			\begin{tikzpicture}
				\node[draw] at (0,6.5) {\huge all touched true};
				\draw[step=1cm,gray,very thin] (-1.9,-1.9) grid (5.9,5.9);
				\draw[fill=blue, thick] (-0.2, 0.2)--(-0.6, 3.1)--(-0.2,4.1)--(1.2,3.4)--(2.8,3.6)--(3.2, 2.6)--(3.4, 2.1)--(1.8,1.3)--(1.2, 0.1)--cycle;

				\draw[ultra thick,->, red] (6.5, 2) --node[above=1mm] {\huge rasterization} (8.5, 2);

				\draw[step=1cm,gray,very thin] (9.1,-1.9) grid (16.9,5.9);
				\draw[fill=blue] (10,0)--(10,5)--(11,5)--(11,4)--(15,4)--(15,1)--(13,1)--(13,0)--cycle;
			\end{tikzpicture}%
		}%
		\caption{Graphical example for rastering a vector (left blue), to a raster (right blue) with weither all-touched False (above), or all touched True (below).}
		\label{fig:alltouched}

	\end{figure}

	\begin{table}[h!]
		\caption{Used levels of costs, the applied numerical equivalent and example layer this cost have been used for.}
		\label{tab:1}
		\centering
		\begin{tabular}{ l  r  l }
			Cost Level & Cost &  Example\\
			\hline
			Prohibited & 500					& National Parks, Buildings \\
			strongly Restricted & 10 	& Bird Reserve \\
			Restricted & 5	& Industrial Areas \\
			No Restriction & 0.5					& Default\\
			Preferential & 0.1					& Power Grid, Motorway Buffers\\
		\end{tabular}
	\end{table}
	The completed list of layers and the processing applied to them, can be found in Supplement S1. \todo{create the Supplement from the processing rules}

	\section{Results}
	In this chapter we want to show the different cost raster, that were created from the very same set of layers,
	but computed for different resolutions.
	From this different raster the cost paths are calculated and compared.
	In the last step the raster with lower resolution were used to calculate in a way, that they shall result in
	similar paths as if the paths were computed from a high resolution raster.
	\subsection{Cost Raster}
	\subsection{Least Cost Paths}
	For each resolution the costs paths were estimated for the rasterization with setting all\_touched False
	and all\_touched True.
	The distance of both paths is calculated with the mean minimum distance.
	For every vertex $P_i$ in the path $L_1$ the minimum between the vertex and the path $L_2$
	is computed and afterwards the minimum distances are averaged (see equation~\ref{eq:1}).
	\begin{equation}
		\label{eq:1}
		d_{mean} = \frac{\sum_{i=1}^{n} d_{min}(p_i, L_2)}{n} \bigg\vert p_i \in L_1
	\end{equation}

	Hence, the mean minimum euclidean distance between the two paths can used to compute
	the similarity.
	As different table~\ref{tab:2} shows the distance between the two paths decreases
	with increasing resolution.
	In addition, this tendency is depicted in figure~\ref{fig:paths_resolution} for the calculated cost paths of 5m and 100 m resolutions.

	At the same time the differences in the aggregated costs stay constant.
	\todo{Interpretation} When normalize the aggregated costs by the resolution.
	Than On one hand it can be seen, that the all\_touched False under estimate the costs and that this tendency scales
	linear with the resolution.
	On the other hand, the all\_touched True least cost over estimates the aggregated costs on a linear scale of
	the resolution.
	Therefore, the difference for the normalized least cost paths is reduced on scale by the resolution.\todo{show map}

	\begin{figure}
		\centering

		\subfloat[\centering resolution of 5 m.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPaths_5m.pdf} }}%
		\qquad
		\subfloat[\centering resolution of 100 m.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPaths_100m.pdf} }}%
		\caption{Figures of the least cost paths. Contrasting the paths for different resolutions. Using OpenStreetMaps as base map.}
		\label{fig:paths_resolution}
	\end{figure}

	\begin{figure}
	\centering

	\subfloat[\centering all touched false.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPaths_al_F.pdf} }}%
	\qquad
	\subfloat[\centering all touched true.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPaths_al_T.pdf} }}%

	\caption{Figures of the least cost paths. Contrasting the changes of the least cost paths for the different results, depending on the parameter all\_touched. Using OpenStreetMaps as base map.}
	\label{fig:paths_alltouched}
\end{figure}

	\begin{table*}[t]
	\caption{Least cost paths as length for the different \acrfull{res} of the rasters, including the \acrfull{mmd} and the \acrfull{agg} costs. From the \acrshort{agg} costs the differences of the \acrshort{agg} costs and the \acrfull{corr} \acrshort{agg} by resolution are given.} 
	\label{tab:2}
	\centering
	\begin{tabular}{ r  r  r  r  r  r  r  r  r}
		\acrshort{res} /m & $length_{\acrshort{al}=\acrshort{f}} /m$ & $length_{\acrshort{al}=\acrshort{t}} /m$ & \acrshort{mmd} /m & \acrshort{agg}  $ cost_{\acrshort{al}=\acrshort{f}}$ & \acrshort{agg}  $ cost_{\acrshort{al}=\acrshort{t}}$ &  $\Delta $ costs & \acrshort{corr} \acrshort{agg} $costs_{\acrshort{al}=\acrshort{f}}$ & \acrshort{corr} \acrshort{agg} $costs_{\acrshort{al}=\acrshort{t}} $ \\
		\hline
		5 & 76136.27	& 78002.00 & 126.04  & 18665.923 & 19616.756 & -850 & 93329.6 &  97584.77 \\
		10 & 75430.10 	& 77936.57 & 277.92 & 8931.245 & 9731.175 & -799.95 & 89312.45 & 97311.75 \\
		50 & 76135.02	& 70619,95 & 1140.01 & 1409.023 & 2300.073 & -891.05 & 70451.15 & 115003.65 \\
		100 & 76283.80	& 74120.73 & 1946.41 & 640.516 & 1572.268 & -931.7 & 64051.6 & 167226.8 \\

	\end{tabular}
	\end{table*}

	Comparing the all\_touched True rasterization and all\_touched False for the same resolution in contrast with the paths of all\_touched False rasterization at the
	different resolutions the later paths are more similar.
	The mean average distance between the 100 m resolution run and the 5 m resolution run is 257.97 m. \todo{Option add 2.5 m and 25 m}

	The similarity of all all\_touched False runs is higher than, the similar between the all\_touched False and all\_touched True runs with the same resolution, except for the highest resolution.

	Hence, in an overall perspective paths of the all\_touched False runs stay in a
	similar region, while paths of the all\_touched True coverage strongly to the paths all\_touched False runs.
	This behaviour is depicted in figure~\ref{fig:paths_alltouched}.
	On a more detailed level, it can be seen, that also the Paths of all\_touched 	converges the all\_touched True path.
	But the extent is smaller.
	The length of the the paths only differs to a maximum of about 10 \%.
	The least path distance for higher resolutions can be lower, because new paths, between regions that are forbidden
	in higher resolution can open.
	On the other side the length of the paths can increase, because with higher resolution more vertices will be used.

	The zonal stat (see table~\ref{table:3}) for a buffer of 100 m (5 m) around the path has been used, to estimate the
	percentage of which costs levels are around the path.
	When using all\_touched True rasterization with higher resolution the tendency is to use a higher percentage of the
	Preferential Level and less the NoRestriction Level.\todo{Interpretation:}
	The ratio of the 100 m buffered least cost path, strongly shifted  to Levels lower costs.
	\todo{Interpretation: Interpretation the Cost Paths are much nearer smaller, than 100 m! No problem because buffering is included in the cost raster.}
	There is no strong tendency for the all\_touched False least cost paths.

	\todo{used regions}

	\setlength{\tabcolsep}{10pt}

	\begin{table*}[t]
		\caption{\acrfull{r} of Category percentages of each least cost path for a buff of 100 m (5 m) around the least cost path.}
		\label{table:3}
		\centering
		\begin{tabular}{ r  r  r r  r r  r r  r r  r r}
			resolution /m & all touched & \multicolumn{2}{c}{ $ r_{Preferential} \% $}  & \multicolumn{2}{c}{ $ r_{No Restriction} \% $ }  & \multicolumn{2}{c}{ $ r_{Restricted} \% $}  & \multicolumn{2}{ c }{ $ r_{strongly Restricted}\% $ } & \multicolumn{2}{c}{ $ r_{Prohibited} \% $ } \\
			\hline
			5 & False &  4.7  &  (5.4) & 58.7 & (58.9) & 8.8 & (8.4) & 0.7 & (0.7) & 27.1 & (26.7)  \\

			10 & False &  19.6 & (33.5) & 68.5 & (64.5)  & 1.0 & (0.8) & 0.8 & (0.3) & 10.1 & (0.9)\\

			50 & False &  20.4 & (33.2) & 68.0 & (66.2)  & 0.9 & (0.1) & 0.7 & (0.0) & 10.1 & (0.5)\\

			100 & False &  21.1 & (30.7) & 69.1 & (68.8)  & 1.1 & (0.0) & 0.7 & (0.0) & 7.9 & (0.4) \\

			\hline

			5 & True  &  18.9 & (28.5) & 67.3 & (66.4) & 1.3 & (1.6) & 1.0 & (0.5) & 11.5 & (3.0) \\	
			10 & True &  18.9 & (33.7) & 66.6 & (63.4)  & 1.6 & (1.4) & 1.4 & (0.6) & 11.5 & (1.0)\\	
			50 & True &  9.1 & (13.0) & 75.7 & (83.0) & 3.9 & (2.0) & 4.2 & (1.6) & 7.1 & (0.4) \\
			100 & True &  7.0 & (10.1) & 73.8 & (81.9)  & 5.5 & (3.9) & 8.5 & (3.6) & 5.2 & (0.4) \\

		\end{tabular}
	\end{table*}

	\todo{Check why values are so different for 5 m all touched False --> simplified line.}

	\subsection{Faster Processing of the Cost Path Algorithm}
	The original least cost path QGIS plug-in uses stops the least aggregation of the costs after the final end point has been found.
	This early stopping might result in sub optimal paths around the end points.
	For further calculation the search for the least cost paths has to be refactored and split the aggregation of the raster for the search of the path.
	To be able to construct a full aggregated cost raster, aggregated cost path in further steps will be computed completely.\todo{really needed}
	For the calculation of the least cost paths.
	The finally least costs paths should between the least cost paths of the lower resolution, with a tendency to be nearer
	The first step to optimize the calculation speed is, only the calculated the least cost paths for this smaller area. 
	\subsubsection{Compare least cost paths, for overlay of both rasterizations}
	\subsubsection{Restrict search to a minimum sized bounding box}
	\subsubsection{Restrict search to a convex hull}
	\subsubsection{Restrict search to reachable points}

	\subsubsection{Compare the different Solutions}
	\subsubsection{Check solution for least cost path between different points}

	\section{Discussion}
	\todo{discuss aggregation with max vs sum}
	\todo{Scaling with multiple endpoints.}
	\todo{discuss other methods. Off algorithmic speed-ups}
	\todo{discuss other technical methods of speed-ups, as precompilation, JIT, vector units, multi-core}
	\section{Related Works}
	\section{Conclusion}

\begin{acks}
	\ldots
\end{acks}

>>>>

\newacronym{wfs}{wfs}{Web Feature Service}
\newacronym{wps}{wps}{Web Processing Service}
\newacronym{res}{res}{resolution}
\newacronym{r}{r}{ratio}
\newacronym{al}{al}{all touched}
\newacronym{t}{t}{True}
\newacronym{f}{f}{False}
\newacronym{agg.}{agg.}{aggregated}
\newacronym{rel}{rel}{relative}
\newacronym{corr}{corr}{corrected}

@@

	\section{Methods}\label{sec:methods}
	\input{methods}

	\section{Results}\label{sec:results}
	\input{results}

	\section{Discussion}\label{sec:discussion}
	As shown need of computation power increases with the power of two with the resolution.
	While the error only scales linear with the resolution.

	As the original least cost path QGIS plug-in the used algorithm stops the least aggregation of the costs after the final end point has been found.
	This early stopping might result in sub optimal paths around the end points.

	As a any buffer around buildings are describes as protected areas, power poles are also included.
	In further works, these should be excluded.

	\todo{discuss aggregation with max vs sum}
	\todo{Scaling with multiple endpoints.}
	\todo{discuss other methods. Off algorithmic speed-ups}
	\todo{discuss other technical methods of speed-ups, as precompilation, JIT, vector units, multi-core}
	\section{Related Works}\label{sec:related-works}

	\section{Conclusion}\label{sec:conclusion}

\begin{acks}
	$\ldots$
\end{acks}
